### 2018-07-17 versions/response.v1.txt
pass thru reviewer response

> Associate Editor
> Editors’ comments to the author:
> All four reviewers agreed that the work is worth publishing and
> provides an important insight into the behavior of the joint
> ML/ancestral states estimation method. However, all reviewers asked
> for what amounts to major revisions to the text and presentation. You
> should especially focus on clarity and accessibility, provide the
> requested examples, and address suggestions on how the results could
> be even further improved.

Outline of revisions:
1.
2.

Some of the reviewers have expressed their opinion that we reduce our emphasis on mathematical development in exchange for a more numerical-optimization-based approach.
Our primary goal was to mathematically prove inconsistency of joint inference, which is what we have done.
As pointed out by reviewer 4, this answers an open question, which has special relevance given the recent work of Sagulenko et al.
We provide some numerical results to confirm our derivations and to provide a comparison.

As we now emphasize (TODO XXX) this work is not the final word on the joint inference question from an applied perspective.
Even if we were to greatly expand the sections doing numerical optimization, these results would only be for the binary symmetric model, which is never applied in practice.
An investigation of these aspects deserves another paper.

> Finally, please pay VERY careful attention to improving your
> literature review (reviewer 4) and making a very clear case about how
> your work IMPROVES and EXTENDS previous results. And, please, do get
> rid of the Farris zone. The name "Farris" should not be attached to
> results that have to do with ML estimation, given the long and
> torturous efforts of cladists to undermine the entire line of inquiry.

We are grateful to reviewer 4 for pointing us to literature of which we were not aware.
We have greatly expanded ...

We have complied with this request regarding terminology, despite it going against current practice.
Indeed, Felsenstein's book has a subsection entitled "The Farris zone" spanning two pages (pp 120-121), a figure entitled "A tree in the Farris zone.", and a table with a caption "Properties of the Felsenstein and Farris zones...".
The term "Farris zone" has almost 7 times as many citations as the term "inverse Felsenstein zone" on Google Scholar.

> Reviewer: 1

> Comments to the Author
> The authors show that when ancestral states are considered as
> parameters, there are zones where topologies and edge-lengths are
> inconsistently estimated. As is evidenced by any ML package, it is
> widely accepted that ancestral states should not be parameters; the
> large number of parameters suggests the approach will be problematic.
> Still, to the best of my knowledge, it has not been shown that
> treating ancestral states as parameters actually leads to poor
> performance, or how poor. Moreover, the authors point to Sagulenko et
> al (2018) - now published in Virus Evolution - as an example
> implementation. I found the paper interesting and I believe a more
> succinct presentation will be of interest.

> A main concern with the paper is that the shaded regions of
> topological inconsistency or poor estimation in Figs 2-4, are only
> subsets of the actual regions where topological inconsistency or poor
> estimation occur; and yet determining the actual regions seems
> feasible. The contrast between Fig 5 and 4 illustrates how far off the
> subsets are in giving the actual regions where estimation is
> problematic. The white region in Fig 5 is the actual region where hat
> w=1 whereas the shaded region of Fig 3 is the subset of the actual
> region that the authors were able to determine through analytical
> arguments. For most readers, out of Fig 3 and Fig 5, only Fig 5 is of
> interest.

As described above, our intent is to mathematically establish inconsistency and poor estimation.
This must be done with analytic arguments rather than optimization.
We could have only shown this bad behavior in a very small area to establish that it happens.
In contrast to this reviewer, we believe that Fig 5 is optional to this line of argument, and we provide it just for interest.
A true investigation into the practical implications of our work will require another paper.

> Getting exact regions is what I think the authors should do. The
> approach used to obtain Fig 5 is generally applicable. Explicit,
> relatively simple, expressions are given in the appendix for long-run
> lnls, l_{tau_1,t*}(tau_2,w,x,y). For any given (x*,y*) on a fine grid
> in Figs 2-4, l_{tau_1,t*}(tau_j,w,x,y), can be maximized using
> numerical algorithms over the three parameters w,x,y. Doing so for
> tau_j=tau_1, tau_2 would determine the actual regions of problematic
> estimation corresponding to Figs 2-4. The only potential difficulty
> that I can see is that numerical algorithms might not converge. This
> can either be avoided by considering a cubic grid of starting (w,x,y),
> or through something like the following. For any given set of
> ancestral character state assignments for the distinct set of 8
> patterns 1111, 1110,... (0000 can treated as the same as 1111), the
> long-run lnl is of the form

> l(x,y,w)=a log(1-x) + b log(1+x) + c log(1-y) + d log(1+y) + e log(1-w) + f log(1+w)

> where a,..f>=0 are constants determined by (x*,y*) and what ancestral
> states have been assigned. The function l(x,y,w) has an explicit
> simple maximizer. Thus the overall maximizer is determined as the
> largest maximized l(x,y,w) over all 4^8 possible ancestral state
> assignments: a finite number of comparisons. The actual number of
> possible ancestral state assignments is quite a bit smaller than 4^8
> (for instance 11 is always the ancestral reconstruction for 1111) and
> the authors have a lot of familiarity with what the ancestral
> reconstructions are. I believe Fig 2 can be replaced with an exact (up
> to grid approximation) region of topological inconsistency and Figs
> 3-5 with heat maps of limiting (x,y,w) for tau_1.

Please see above for our response.
%das: I added some stuff below to maybe alleviate some confusion and to show why what this reviewer thinks we should do won't work, but not sure how much is necessary.
It is true that the form given above for l(x,y,w) has an explicit and simple maximizer.
However, the long-run log-likelihood is more complicated than this function, as seen in equation (9).
For both topologies, we have a number of equations of this form that we would need to maximize in a grid search.
The function of the form in l(x,y,w) above was our approximation to the upper and lower bounds of the log-likelihood, and was our choice for these bounds precisely because it has an explicit and simple maximizer.
Doing a grid search over all ancestral states would not be trivial, and we would need to perform numerical optimization for all possible ancestral states to compute the maximum.
Moreover, not assuming x_1=x_2 and y_1=y_2 (see below) results in an explosion of functions to perform a grid search over, with none of these functions being guaranteed concave.
The analytic bounds we have provided were devised specifically to sidestep this nonconcavity.

> The other main concern that I have is with the restriction that
> t=(x,y,x,y,w) in estimation rather than t=(t_1,...,t_5). On the one
> hand, usually one expects that incorporating knowledge about the true
> underlying tree (t_1=t_3 & t_2=t_4 in the present case) should be
> helpful for a good estimation procedure. Thus the fact that treating
> ancestral states as parameters gives inconsistent topological
> estimation in this setting demonstrates that the approach is
> problematic. On the other hand, allowing t=(t_1,...,t_5) is what would
> be done in any practical implementation and it seems quite possible to
> me that results would be radically different in this case. For
> instance, from Fig 2, the generating tree with all equal edge-lengths
> (y*=x*=0.2) provides a case where topology tau_2 is inconsistently
> estimated with the restriction that estimated t=(x,y,x,y,w). Because
> of the symmetry of the problem, I am doubtful y*=x*=0.2 would be a
> point of inconsistency if t=(t_1,...,t_5) were unrestricted in
> estimation. The issue deserves discussion because I don't think it is
> transparent.

TODO: make the argument more clear in the main text.
%das: BIIIIG update
We thank this reviewer for highlighting an omission in our argument.
We had initially simplified the results by making the erroneous assumption that t_1=t_3 and t_2=t_4.
It can be shown that this assumption ignores potential feasible parameter sets, and we have removed it.
This requires a change in the arguments in the theorems as well, though the conclusions are similar.

> I don't think the restricted case of p8-9 will be of interest to
> readers. Only the general case.

%EM I think we could move this to the appendix to throw this reviewer a bone.
%das: another big update
We have removed the restricted case as it does not reflect how inference is performed and was only included as an expository example.
Given the change to the theorem arguments, we do not believe this case is necessary for highlighting inconsistency.

> Minor Comments:

> -l78-88: Calculations are more natural with `fidelities' and this
> might remain in the appendix but it is much more conventional in the
> literature for x,y,w to be edge-lengths or probabilities of a
> different character states at the tips of the edge. It takes quite a
> bit of getting used to realize that (0,0) corresponds to infinite
> edge-lengths and it seems unnecessary with a little work on the part
> of the authors to do the conversions that will make the figs more
> natural for readers.

Thank you for this suggestion.
We have implemented it in all the plots and have been careful to clarify that "fidelities" do not correspond to "lengths" in the text, deferring to "lengths" where appropriate.

> Reviewer: 2

> Comments to the Author
> In this manuscript  the authors show that joint inference of phylogeny
> and ancestral characters is not consistent in general by providing a
> counter-example, in which as the number of sites grows, they show that
> the wrong tree topology can have a higher likelihood than the correct
> topology under the joint inference model.
> The result is important for the community to understand, in particular
> because TreeTime (the joint inference method) is used in high-profile
> analyses on Ebola, influenza and other viruses on the nextflu /
> nextstrain websites. These are beautiful and accessible sites aiming
> to make analyses on viruses open to exploration by the community, but
> this comes at the cost of suppressing the details of the analysis. If
> indeed there are deep statistical issues or limitations it is
> important that the community knows what these are and what their
> implications are.

> Accordingly my comments are focussed on where I think the manuscript
> could be more clear (and where I did not quite follow the logic). My
> major suggestion would be to go even further towards demonstrating in
> a simple example how the joint inference can be wrong – for example,
> providing a toy multiple sequence alignment or even one of some
> influenza HA sequences – you could find 2 sets of 4 closely-related
> sequences that have the Farris and Felsenstein topologies, maybe? And
> demonstrate that the inference gives the wrong result by explicitly
> computing the likelihood.

We agree that understanding the implications of our derivation is a very interesting avenue for future work.
However, to do such a project justice would require a separate paper.
We are sure that once our paper demonstrates this inconsistency, others will quickly follow to write such a paper.

> Also, laying out the theory a little bit more step-by-step so that a
> general quantitative reader doesn't struggle too much to follow would
> help (suggestions below).

Thank you very much for reading the derivations and for your suggestions. As noted inline we have revised the text accordingly.

> I know that it's not feasible to explicitly compute (analytically)
> these likelihoods for bigger trees. But one of the main motivations
> and applications for tools like TreeTime is surely that the number of
> sequences grows (not the number of sites, which is fixed by the viral
> sequence length in the case of viruses). What happens when the number
> of tips grows? Are there more opportunities for the kind of issue that
> gives the wrong value for w? Many influenza trees have many 0 (or near
> 0) internal branch lengths, and it would be interesting to understand
> the implications for the ancestral state reconstructions (which in
> turn are used to infer spatial spread, transmission etc).

%EM David-- does it seem like we could extend these results to bigger trees containing a 4 taxon tree?
%das: I wasn't able to come up with anything interesting for this.

> Some detailed comments and questions:
> – line 85 some cut/paste or sentence issue

TODO: not sure what this is?
%das: this reviewer was confused about the phrase "generating probabilities," thinking we were generating the probabilities and not that they were the probabilities of generating a particular data configuration.
We have added the definition of a "generating probability" to ease confusion in this sentence.

> - line 97 below (2) define A^p

We have clarified this with an explicit definition of A^p where we introduce A, at the beginning of the section on Phylogenetic maximum likelihood.

> - l106: why is the functional form of (3) determined by inequalites;
> what are they?

We have made this more precise and directly referenced the inequalities.

> - l 118-124: clarify are tau1 and tau2 Farris and Felsenstein?

That is correct, and we've added parenthetical notes to reminde the reader, thanks.

> L 132 clarify “equivalent but different notion” – different from what?

We have reformulated this sentence to avoid the comparison.

> - line 146 “seeing no change along a branch more likely increases the
> likelihood by introducing a term of (1 + θ) instead of (1 − θ)” – I
> don't understand this

TODO: this is changed, but I don't see this being a lot clearer.
%das: I believe this is clear now; lmk
We have clarified the intuition in this sentence.

> line 185++ and figure5: true value of w?

We have added clarifying sentence:
"As data were generated with parameters $\{x^*, y^*, x^*, y^*, y^*\}$, the true value for $w$ is $y^*$."

> L 334 missing “the”

Fixed via simpler language.

> L332++ I got confused about the complements and factors of 2, though I
> see the point that the site split has symmetry. As it's fundamental to
> the rest of the work, a more clear statement of where the 2 comes from
> in the second eq below 358 would be very helpful.

We have now made this much more explicit with three additional display equations spelling out the intermediate steps.

> L354 why does psi take the complement of site patterns? Just above it
> is defined as a random variable psi (Y) where Y is a site pattern

This wasn't as clear as it should have been. We have revised to:
"The mapping $\patternToSplit$ operates by returning the complement of a site pattern to obtain a site split in $\mathcal{P}(\siteSplitSet)$ if it is not in $\mathcal{P}(\siteSplitSet)$."

> What are the elements of the sets eta(tau, t)?

We removed the ordered list of sets as it did not add to any understanding of the concept and now refer to each of these sets and their relevance directly.

> Why is xi surjective? I don't understand the equation before line 359,
> and then I don't understand what “such a choice” is in line 359

We have made this explicit now, with two additional intermediate steps and an explicit explanation in terms of collapsing terms.

> line 364 define p(e)

Thanks for reminding us to define p(e), which is now described as "where $p(e)$ is the probability of a character change along edge $e$."

> line 366  I don't understand (8) but perhaps that's just background I
> don't have – a quick explanation or reference would be handy
> What's a 'generating probability' ? Is that the fidelity? Or a
> generator?

We have updated the text to explicitly define "generating probability." Background for equation (8) can be found in Section 8.6 in the Semple and Steel text. This is referenced at the beginning of the section. We do not believe an exhaustive explanation would provide any benefit to the reader that they could not get from referring to the text.

> Reviewer: 3

> Comments to the Author
> Shaw and Matsen show that a joint maximum-likelihood inference of
> phylogenies and ancestral states is not consistent when branch lengths
> are long. We have recently proposed such a joint estimation for
> sequences that recently diverged and are hence similar (Sagulenko et
> al). The results presented here are anticipated in Sagulenko et al
> "For more divergent sequences, however, subleading states of internal
> nodes make a substantial contribution and the iterative optimization
> will underestimate the branch lengths."
> and
> "The iterative optimization procedures are not expected to be accurate
> for trees were many sites are saturated. In scenarios with extensive
> uncertainty of ancestral states and tree topology, convergence of the
> iterative steps cannot be guaranteed."

> In this light, it should be pointed out that joint estimation
> procedure was explicitly suggested as an approximation and hence never
> expected to be consistent in a strict sense. Furthermore, it was
> suggested as an approximation when branch lengths are short -- the
> opposite of the problematic regions identified here.
> But systematic analysis of the behavior of this approximation is
> useful and I believe the work is correct (but haven't checked the
> proofs). The work by Shaw and Matsen suggests the radius of
> convergence of this procedure is somewhere around branch length of
> order 10-20%, which is consistent with results reported by Sagulenko
> et al.

Thank you for your thoughts, and we have worked hard to not oversell our results.
The TreeTime paper does present joint optimization as an approximate approach, and described that it is for short branch lengths.
However, in our assessment the word "approximate" doesn't completely describe a situation in which there is significant bias.
The TreeTime paper says that its use is appropriate when "ancestral sequence inference has little ambiguity" but even in these cases the bias in branch lengths persists.

TODO we should make some changes to at least say that we've further clarified the implications of our results.
%das: how much more should we say than this sentence (which we already say)? We already quote them as saying they should only use it when branches are short.
%Although the software described in \citet{Sagulenko2017-jo} fits on a user-supplied topology and the authors explicitly warn that the approximation is for the case where ``branch lengths are short and only a minority of sites change on a given branch,'' their work motivates understanding the general properties of such joint inference.
%and
%In all cases, these regions of inconsistency arise when the branches of the generating trees are ``long,'' that is, when branch fidelities tend to be small.
%This inconsistency in the case of long branches concurs with some empirical findings in \citet{Sagulenko2017-jo}, namely their Figures~2 and 3.

> At the same time, these results are not particularly surprising.
> Assigning ML ancestral states will always result in shorter branches.
> For short branches, this correction is linear in branch length, for
> longer branches higher order effects start to contribute. If the
> typical branch length around a bifurcating node is x<<1, a fraction ~x
> of the sites mutated along one of the three adjacent branches, while a
> fraction ~x^2 mutated on two branches and not the third. The latter
> case is ignored in this joint inference procedure. Not having to sum
> over these subleading states is what makes joint inference faster, but
> such truncation will first lead to inaccuracies and at longer
> distances to qualitatively different solutions (like Taylor expanding
> 1/(1-x) ).
>
> There are other obvious cases where joint inference will fail. For
> sequences with a heavily biased in nucleotide composition (say biased
> towards A). For sufficiently long terminal branches, all ancestral
> states will be estimated as all 'A' and subsequent branch length
> estimate will shrink all internal branches to zero. This happens for
> the very same reason that the most likely outcome of multiple flips of
> a biased coin is all heads -- a very atypical yet most likely outcome.
>
> Currently, the work largely shows that areas of inconsistencies exist.
> It would be useful if the authors could present Taylor expansions for
> the ML branch length estimates around x=y=w=1. This would more
> explicitly show the inaccuracies accumulate with increasing branch
> length and how the inaccuracies depend on asymmetries in the tree. To
> me, this would be much more useful than to know that joint inference
> will fail if branches are long.

This is an interesting suggestion but is different than our goal, which was to establish inconsistency.
Moreover, in the revised manuscript we see cases where assigning ML ancestral states results in longer branches, not shorter ones (Lemma 3 in the updated paper).
We believe the inaccuracies have more to do with the specific choice of which ancestral state is maximal than asymmetries in the tree.
Asymmetries can drive these particular choices, but Fig. 3 in the revised manuscript shows that almost half of the parameter space results in branch lengths being estimated as zero.
We do not think providing an argument about accumulating inaccuracies would elucidate why such starkly-defined regions of inconsistency exist.
For reasonable parameter settings, one could investigate the error accumulation given fixed ancestral states, but this was not our goal and will require another paper.

TODO describe why Taylor isn't a straightforward thing.
TODO refer to idea of applied paper again?

> Specific comments:
>  - when y=x=w, the two topologies are indistinguishable other than for
>  the labeling of the leaves. What happens in Fig 2 on the diagonal?
>  Does it degenerate into a star?

It is true when y=x=w we are unable to distinguish the topologies apart from the labeling, though the labeling is what distinguishes the topologies.
As shown in Goldman (1990), with all branch lengths fixed and equal the joint inference procedure degenerates into parsimony, which has well known inconsistency issues.
We have added discussion of this point.

>  - line 50: overestimates branch lengths? It overestimates fidelity ->
>  underestimates branch length.  As expected.  Again line 249. Please
>  check. The back and forth between fidelity and branch length can be
>  confusing.

Thank you for bringing this to our attention.
We have tried to use branch lengths in as many instances as we could in the discussions, using fidelities only in the computations as they allow for great simplifications.

>  - Fig 5: it is surprising that \hat{w} seems completely flat as a
>  function of x* until it jumps discontinuously to 0. Why not include
>  the analytical bound in Fig 5?

The discontinuity is a reflection of the change in ancestral states when the generating parameters change.
As the true generating parameters change, we become more likely to choose a different ancestral state as most likely, resulting in an entirely different objective function to be optimized.
The analytical bound is a coarse estimate of where inconsistency arises.
In the revised manuscript, we fix the maximal ancestral state to simplify the argument, and comparing the analytical with the empirical would not provide much added benefit to the analysis.

>  - It would be nice if the abstract pointed out that these
>  inconsistencies arise when branch lengths are long and that the
>  approximate joint maximization procedure was suggested for precisely
>  the opposite limit. Similar at the beginning of the discussion, I
>  think it would be more appropriate if this would say smth along the
>  lines "Sagulenko et al suggested an approximation for short
>  branches... we show that this doesn't work for long branches..."

TODO we should do some change, but make it clear that it's consistently biased even for short branches.
%das: I added this line:
We have added the line:
\citet{Sagulenko2017-jo} explicitly warn their approximation is only valid for ``short'' branches, though Fig.~\ref{fig:bl-general-bias} shows even when branches are short there is a consistent bias.

We believe it's important not to mischaracterize the suggested use of joint inference, though we would like to emphasize to the reader that problems are apparent in large regions of the parameter space.
We think the current discussion treats this optimization procedure as fairly as possible.

> Richard Neher

> Reviewer: 4

> Comments to the Author
> Review of Shaw, D. and Matsen, F.  "Joint maximum-likelihood of
> phylogenies and ancestral states is not consistent"

> Shaw and Matsen demonstrate that the approach of jointly estimating
> ancestral states and tree topologies (or branch lengths on a fixed
> topology) is not statistically consistent.  I believe that their
> results are correct, and that their result is important to
> disseminate.  The paper is clearly written, given the complexity of
> the material, and I have no reservation about recommending publication
> virtually as is.

> However, although my original intent was to carefully work through the
> math (including the derivations and proofs in the appendix), I have
> not found the time to do so, and because of the lateness of my review
> at this point, I will provide just a few comments in the hope that
> they will be useful.  Hopefully, other reviewers were able to check
> the correctness of the mathematical results more carefully.

Thank you for your assessment.
We have worked hard in our initial submission and in our revision to write out the argument in great (indeed, excruciating) detail, and we would appreciate it if you could take a look at the revision.

> The only real criticism I have of the paper is that the authors could
> have done a better job in connecting their new work to previous
> results.  For example, Goldman (1990, Syst. Biol. 39:345-361) showed
> that joint estimation of topology and ancestral states by maximum
> likelihood was equivalent to maximum parsimony analysis for the same
> 2-state model considered by Shaw and Matsen when all branches are
> forced to have the same length.  They noted the statistical
> inconsistency of this approach, which was necessarily true because
> parsimony was already known to be inconsistent (the famous result of
> Felsenstein 1978).  Shaw and Matsen extend this result to the case of
> unequal branch lengths, which while not particularly surprising, is
> nonetheless an important extension.  The authors should also cite key
> relevant earlier work on the identifiability ot the tree topology with
> "classical" maximum likelihood phylogenetic estimation (Allman et al
> 2008, Adv Appl Prob 40:229–249; Chai and Housworth 2011, Syst Biol
> 60:713–718) in addition to crediting Roychoudhury 2015 for
> establishing ML consistency.

We are very grateful to this reviewer for pointing out the Goldman paper.
We were not aware of this paper and now have a paragraph describing it and comparing to our work.
It's also a good idea to cite additional papers about identifiability, and we have done so.

%EM David, heh, can you point me to the place where the equal branch length result was done?
%EM Also, I think you said that he says that the varying branch length thing is an open question-- it would be great to quote that.
%das: update I added some of the below. Let me know what you think. And re: Allman/Chai, I just added them in with RoyChoudhury; is that ok? They are all just examples of conditions where phylogenetics is consistent it seems.

%das: below are some notes and quotes. I will work on paring these down and adding some to the text, maybe with a little more diplomacy.
%The equal branch length example occurs in Example 2, starting halfway down the second column of page 352 in Goldman (1990) and continuing until the top of the second column on page 354.
%Here he only talks about how parsimony is inconsistent because it may result in equally parsimonious trees (this is the last paragraph of the example section).
%It's not until the bottom of page 356 that he punts the inconsistency to Felsenstein and Hendy/Penny, and the bottom of page 357 where he wonders how small branch lengths have to be for parsimony to be consistent in general.

---

Goldman (1990) concerns estimating both structural (e.g., topology, branch lengths) and incidental (e.g., ancestral states) parameters simultaneously, though the only worked example is in the case of a topology with equal branch lengths.
Goldman shows that for certain special cases, "maximum relative likelihood," i.e., performing maximum likelihood over both topology and ancestral states, is equivalent to parsimony, and thus inconsistent.
He says
"Consistent estimates of incidental parameters are unobtainable, as it is not possible to increase indefinitely the number of observations (data) pertinent to each."
However, he also says "results may be unpredictable" and "estimates of the structural parameters (e.g., the tree T) may also be inconsistent."
He does not discuss the inconsistency of maximum relative likelihood in general, and proceeds to show that parsimony analysis is a specific instance of maximum relative likelihood, implying maximum relative likelihood of topology given equal branch lengths cannot be consistent in general, but stops there.
We show cases where the estimates of the topology will be inconsistent if we are estimating both topology and branch lengths, and, moreover, even if we know the correct topology, we show cases where branch lengths will be inconsistent.
Finally, just prior to his conclusion, he discusses when parsimony will give the same answer as ML, concluding that the question is ill-posed since parsimony estimates different parameters than ML, i.e., it assumes equal branch lengths.
He states that if the expected amount of change along a branch is "very small" then parsimony will recover the same tree as ML (under a Poisson process model).
We are able to quantify how small is "very small" in the case of general joint ML---estimating topology and branch lengths together---and not just parsimony in particular.

quote:
"The second approach is to estimate the random variables as though they were parameters of the model. However, in this case they will be incidental parameters: as the amount of data (i.e., the number of characters) increases, the number of parameters also increases. For each additional character (labeled a, for instance), there are additional data xa, consisting of the states for all n species, and additional parameters Ya, consisting of the states for all the internal nodes {NI: I = 1,2, . . , m}. Although this method is the only one possible if we wish to estimate the internal node character states, the results may be unpredictable. As noted above, these estimates cannot be consistent. Further, the estimates of the structural parameters (e.g., the tree T) may also be inconsistent. Indeed, under certain models, this approach renders inferences virtually impossible (Cavalli-Sforza et al., 1964; Cavalli-Sforza and Edwards, 1966; Kendall and Stuart, 1979). Examples of both approaches are investigated below."

---

> Finally, let me just add that I hate the term "Farris Zone".  The
> "Felsenstein Zone" was a name applied to Felsenstein's important
> demonstration of a tree and branch-length configuration where
> parsimony would be an inconsistent estimator of the phylogeny.  On the
> contrary, Farris had nothing to do with the "Farris zone"; it was
> simply a name given by Mark Siddall to a different configuration in
> which parsimony outperforms likelihood, thinking incorrectly that
> likelihood was inconsistent in this zone.  Swofford et al. (2001,
> cited by the authors) referred to this zone as the "inverse
> Felsenstein zone", which I think is a better name for it.

As described at the beginning, we have made this change, thanks.