### 2019-01-11 cover-letter/supervisor.tex
update cover letter

\documentclass[letterpaper,10pt]{article}

\usepackage{cite}
\usepackage{enumerate}
\usepackage{url}

\hyphenation{Ge-nome Ge-nomes through-put re-comb-in-ation}

\input{FHletterdefs}

\renewcommand\FHtoname{}
\renewcommand\FHtoaddress{}

\renewcommand\FHopening{Dear MBE Editorial Board,}

\begin{document}

Please consider our revised manuscript entitled ``Joint maximum-likelihood of phylogeny and ancestral states is not consistent'' for publication in \textit{Molecular Biology and Evolution}.
In this paper, my co-authors, David Shaw and Vu Dinh, along with myself, investigate the statistical properties of joint maximum-likelihood estimation of trees and ancestral sequences.
We were motivated to do so by recent work from Richard Neher's group (\url{http://dx.doi.org/10.1093/ve/vex042}) that employs such joint optimization as an efficient alternative to the standard Felsenstein marginalization.

We show that such joint inference is not consistent, both for topology inference and branch length optimization.
For a fixed topology, almost the entirety of the parameter space yields inconsistent branch length estimates.
Additionally, there are sizeable regions of parameter space in which this joint inference gives the wrong topology in the limit of large data.
This is not a case of long branch attraction.
Indeed, the inconsistency in Felsenstein 1978 requires a combination of long and short branches, whereas our inconsistency arises in roughly half of the parameter space.

Although the proofs in our paper require detailed examination of likelihood functions, we have written the paper to be broadly accessible.
The paper has two illustrations showing regions of inconsistency for topology and branch length inconsistency, respectively, and provides intuitive summaries of the results in the main text.
A complete set of proofs appears in the appendix.

We feel that it is important for this paper to get a critical review by a mathematically-oriented phylogenetics expert.
We suggest:
\begin{itemize}
\setlength\itemsep{1.4pt}
\item Elizabeth Allman \texttt{esallman@alaska.edu}
\item Laura Kubatko \texttt{lkubatko@stat.osu.edu}
\item Mike Steel \texttt{mike.steel@canterbury.ac.nz} (although Mike is somewhat conflicted as my postdoc advisor, he would do an excellent job reviewing this paper)
\item Ed Susko \texttt{susko@mathstat.dal.ca}
\item Richard Neher \texttt{richard.neher@unibas.ch}
\end{itemize}

Thank you for your consideration.
\end{document}