### 2019-03-19 versions/response.v2.txt
Adding response template and old version of ms

> The reviewers have universally viewed the revision as much improved.
> That said, 3/4 maintain that further changes are necessary prior to
> publication. It is our view that the remaining criticisms are "minor",
> and should you address them thoroughly and convincingly, we expect
> that final acceptance of the revision can be done editorially.

XXX

> Reviewers' comments:
> Reviewer: 1

> Comments to the Author
> As with the previous version, I believe the topic matter will be of
> interest. For me, however, the work is incomplete in important ways
> that I believe can be relatively easily addressed. The authors went
> beyond my expectations in calculating explicit solutions for the
> limiting edge-lengths (Fig S2) when tau1 is the putative topology of
> estimation. This allows them to draw conclusions about edge-length
> inconsistency but there is no definitive conclusion about `strict'
> topological inconsistency (in the sense of converging upon the wrong
> tree rather than a multifurcating one). From the statements near l238
> it seems the authors have completed some preliminary investigation on
> strict topological inconsistency and find no evidence for it at
> present. Assuming that remains true, the findings of the current ms
> are of a different nature than the previous version where strict
> topological inconsistency was found. In settings with a limiting
> multifurcating topology, because it is multifurcating, for finite but
> large sequence-lengths, measures of uncertainty (eg. bootstrap
> support) will suggest that there is substantial uncertainty in
> whatever topology gets estimated (this probably deserves discussion).
> The joint optimization approach thus lacks the power to estimate the
> correct topology where ML would but is not being positively mislead.
> Assuming the lack of strict topological inconsistency holds, the
> results for the method are not as negative as they were in the
> previous version.

TODO: add discussion of bootstrap

> Getting complete results about topological consistency is feasible.
> The arguments that led to the closed form solution for the limiting
> edge-length estimates for tau_1 can be repeated similarly for tau2.
> This would allow, for any given x* & y*, comparison of the limiting
> lnl per site for tau1 vs tau2. Whichever topology has the larger lnl
> is the long-run estimated topology. In cases where hat w=1 for both
> topologies, conclusions are different but definite. Thus it seems a
> relatively straightforward exercise given the work the authors have
> already done (I hope I am not missing a subtle difficulty) to have a
> computer determine for a grid of (x*,y*) whether (a) hat w=1 exactly
> for both topologies, (b) the difference in limiting lnl over
> topologies is either positive or negative or (c) hat w neq 1 for at
> least one of the topologies and the difference in limiting lnl is 0 up
> to machine precision. I would emphasize that the cases (a) and (b)
> constitute rigorous proof. The cases (c) likely are rare and
> correspond to boundary cases, so the boundary may have some slight
> uncertainty. There may be some aesthetic appeal to getting analytical
> results rather than using a computer but being able to draw clear
> conclusions about strict topological inconsistency is important.

DISCUSS: we have no evidence of topological inconsistency

> The language used to refer to the approach and references to the
> literature are awkward. The references to profile likelihood work
> suggest that there is a theoretical basis for the approach (that paper
> suggests consistency) but the settings considered in such work are
> quite different, I believe. Integrated likelihood usually refers to
> something different too. Finally, as Reviewer 3 makes clear the only
> current implementation of the approach was intended as an
> approximation to ML rather than as ML. The central issue is whether
> treating unobserved data (ancestral states here) as parameters
> sometimes works; the EM algorithm were developed with the idea that
> this is not what should be done. In any case, references directly tied
> to that issue would be more relevant. Additional suggestions include:
> 1) Make a clear statement in the intro about why the usual likelihood
> is considered the right likelihood: it is the probability of the
> *observed* data. 2) The point that large numbers of parameters need
> not defeat likelihood methods is valid but it should be made clear
> when referencing such work that the setting is different.

> -l21: Some rewording is needed. The joint method will, as n gets
> large, continue to estimate a well-resolved topology with positive
> probability; albeit with small middle edge-length.

> -l131: A reader might be able to avoid the appendix (which is
> difficult to wade into) if an example zeta in a four taxon case was
> given as illustration.

> -l150: I initially found this paragraph confusing. Table S1 is a
> result that you are using as intuition, so you can't be giving
> intuition for the entire result (otherwise it gets circular). I think
> you are trying to give intuition for the `everywhere except a set of
> measure' portion of the result rather than entire result. Perhaps `the
> proof that hat t neq t* except on a set of measure 0 can be seen
> intuitively...'.

> -l154 and (should check elsewhere): `the limiting value of the
> estimator' rather than `the estimator' is what is being referred to
> here.

> -l186: which `likelihood terms' is not that clear to me: `each site
> results in multiple factors, only one of which depends on w with
> dependence being through (1+w) or (1-w) depending ...'

> -l189: `depends on the values of other branch parameters' or some
> other rewording. I don't think it is that they are unknown. It is that
> they need optimization, so you can't optimize w independently of them.

> -l201, Figs 2 and 3: I don't understand why direct numerical
> optimization was needed. You have the exact expressions for hat w in
> Fig S2. So you could have plotted them as Fig 2. A similar comment
> applies to Fig 3. Are you trying to convince the reader that your
> proofs have no errors? That seems more an issue for supp material.

> -l252-259: I don't think this is an example of profile likelihood as
> considered in Murphy and Van der Vaart.

> -l264: Do you mean that `consistency can hold even when the numbers of
> parameters tend to infinity'? The statement `not a necessary
> condition' just means that inconsistency can arise in different ways
> (eg. lack of identifiability).

> Reviewer: 2

> Comments to the Author
> Shaw et al provided a revised version of the manuscript that I
> reviewed about a year ago showing that joint estimation of ancestral
> states and branch length is in general not consistent. The authors now
> map out the regions in which topologies are incorrectly estimated and
> include a plot that shows the bias of the internal branch length for
> different tree configurations. I think this is a useful piece of work.
> But some parts are completely unsurprising, while some empirical
> results seem inaccurate or misleading.

> The most important results/claims by Shaw et al are:

> 1) "jointly estimating phylogenetic parameters along with ancestral
> states is not consistent in general"... "only correct on a set of
> measure 0".
> It is clearly true that branch length will be biased. But the fact
> that joint estimation will not recover the true branch lengths is not
> surprising but perfectly expected. Joint estimation ignores subleading
> configurations of internal states. Since the subleading terms
> typically require more state changes, branch length estimates ignoring
> them are biased. In a similar way a Taylor expansion is almost nowhere
> "correct", but still a useful and insightful approximation (a better
> analogy is ignoring looped diagram in a diagrammatic expansion.).

> 2) In some parameter regions the internal branch of the four taxon
> tree collapses to 0.
> Again, this is true and identifying regions in parameter space where
> the internal branch collapses to zero is the most interesting part of
> the manuscript in my opinion. Unfortunately, the manuscript doesn't
> give much insight as to why the length of the internal branch shrinks
> to zero in some regions of parameter space. There is, however, a
> rather intuitive explanation for this. In the inv Fels topology, the
> internal branch collapses to 0 if
>
> p_x^2 > c p_y
>
> (the prefactor c depends on specifics like alphabet size etc). And
> this has an intuitive interpretation: if p_x^2 > p_y, it is more
> likely to observe a mutation twice on two branches of length p_x than
> once on a branch of length p_y. As a result, the joint inference will
> push p_w to zero and assign all state changes on the terminal branches
> of length p_x. But for short branches, this corresponding sector of
> parameter space is a small (tending to zero for a fixed p_x/p_y with
> p_x \to 0).
>
> My empirical tests of the approximate joint inference scheme suggest
> this region is much smaller than the one identified by Shaw et al in
> Fig 3.

> 3) "[The bias is] of the same order as the branch lengths themselves."
> This is incorrect or misleading, depending on how you read it. In
> typically cases, the error made by this procedure should be quadratic
> in branch length and hence of higher order than the branch length
> itself. All empirical test we have done via numerically optimizing
> branch lengths suggests this is true for typical trees and that the
> errors made are small provided ALL adjacent branch lengths are short.
> That said, relative error of one branch can be large if adjacant
> branches are long such as the case above. In this case the error is
> quadratic in the length of the longer adjacent branches rather than
> the branch itself.

> Reviewer: 3

> Comments to the Author
> This a resubmission of an earlier version of the paper, for which I
> was reviewer #4.  The other reviewers of the previous submission did
> an excellent job in pointing out minor problems and making suggestions
> for improvement.  It seems that Shaw et al. have addressed all of
> these suggestions in exemplary fashion, and  think the result is a
> much stronger paper.  I especially appreciate the authors' efforts to
> make the mathematical derivations and proofs more accessible to
> general readers (like myself).

> I often write long reviews but given the careful attention to detail
> in responding to the previous reviews of an already good paper, I have
> nothing further to add.

> Reviewer: 4

> Comments to the Author
> The authors have made some substantial changes addressing the concerns
> of the last round and I think that has improved the clarity and in
> some instances also strengthened the results.
> Remaining comments are mostly about the notation which I still find a
> bit hard to follow.
> Why are site and ancestral node splits carrying this distinction of
> whether the m'th taxon (or node) has a 1?

> Perhaps it's a latexdiff leftover but eq 6 in line 151 seems to have a
> line with just a ; \xi )

> Line 593 p 25 I don't understand "the mapping \psi operates by
> returning the tips labeled as 1 in a site pattern to obtain a site
> split in P(Y) if the set of tips labeled 1 is not in P(Y)" and this
> mapping seems fundamental. Perhaps you mean to refer to the m'th taxon
> (or node)..?

> The symmetric model is assumed strongly in the first part of the
> appendix, with the factor of 2; my guess is that the results hold in
> asymmetric models but the notation is harder, is that correct?

> What's q in equation 7 in the appendix (line 655) ?

> More substantitvely, the limit in (8) seems not established because as
> n -> \infty so would q and n_j, so these can't be treated as fixed .

> page 9 - at this point in the paper, unless the reader has already
> read the appendix they don't know what the estimator means or what the
> p_ytilde j are - so this text doesn't make sense at this point.

> page 30 line 667-8 what are S and Y and what makes a path? I gather
> from the math that you mean paths in the tree, since the sum is over
> ancestral states and there's a product of probabilities of changes on
> branches but (perhaps my lack of background) the form of (9) isn't
> obvious I think because I am not exactly clear on what the notation
> means.