### 2019-03-25 versions/response.v2.txt
I'm happy.

> The reviewers have universally viewed the revision as much improved.
> That said, 3/4 maintain that further changes are necessary prior to
> publication. It is our view that the remaining criticisms are "minor",
> and should you address them thoroughly and convincingly, we expect
> that final acceptance of the revision can be done editorially.

Thank you for handling the manuscript, and please thank the reviewers for their helpful comments. As stated by Reviewer 3, we believe that the review process has significantly improved the manuscript.

> Reviewers' comments:
> Reviewer: 1

> Comments to the Author
> As with the previous version, I believe the topic matter will be of
> interest. For me, however, the work is incomplete in important ways
> that I believe can be relatively easily addressed. The authors went
> beyond my expectations in calculating explicit solutions for the
> limiting edge-lengths (Fig S2) when tau1 is the putative topology of
> estimation. This allows them to draw conclusions about edge-length
> inconsistency but there is no definitive conclusion about `strict'
> topological inconsistency (in the sense of converging upon the wrong
> tree rather than a multifurcating one). From the statements near l238
> it seems the authors have completed some preliminary investigation on
> strict topological inconsistency and find no evidence for it at
> present. Assuming that remains true, the findings of the current ms
> are of a different nature than the previous version where strict
> topological inconsistency was found. In settings with a limiting
> multifurcating topology, because it is multifurcating, for finite but
> large sequence-lengths, measures of uncertainty (eg. bootstrap
> support) will suggest that there is substantial uncertainty in
> whatever topology gets estimated (this probably deserves discussion).
> The joint optimization approach thus lacks the power to estimate the
> correct topology where ML would but is not being positively mislead.
> Assuming the lack of strict topological inconsistency holds, the
> results for the method are not as negative as they were in the
> previous version.

The reviewer is correct that this story is slightly different than the one in the previous version of the paper. Regarding the bootstrap, we remind the reviewer that we are interested in the case of "infinite data," and thus the bootstrap doesn't apply. We have also taken out the speculation about higher likelihoods on the wrong topology.

> Getting complete results about topological consistency is feasible.
> The arguments that led to the closed form solution for the limiting
> edge-length estimates for tau_1 can be repeated similarly for tau2.
> This would allow, for any given x* & y*, comparison of the limiting
> lnl per site for tau1 vs tau2. Whichever topology has the larger lnl
> is the long-run estimated topology. In cases where hat w=1 for both
> topologies, conclusions are different but definite. Thus it seems a
> relatively straightforward exercise given the work the authors have
> already done (I hope I am not missing a subtle difficulty) to have a
> computer determine for a grid of (x*,y*) whether (a) hat w=1 exactly
> for both topologies, (b) the difference in limiting lnl over
> topologies is either positive or negative or (c) hat w neq 1 for at
> least one of the topologies and the difference in limiting lnl is 0 up
> to machine precision. I would emphasize that the cases (a) and (b)
> constitute rigorous proof. The cases (c) likely are rare and
> correspond to boundary cases, so the boundary may have some slight
> uncertainty. There may be some aesthetic appeal to getting analytical
> results rather than using a computer but being able to draw clear
> conclusions about strict topological inconsistency is important.

In our revised work we have found no evidence of topological inconsistency, and as mentioned above we have removed such speculation from the paper. "Machine precision" is not involved: w exactly equal to 1. That's its optimal value.

> The language used to refer to the approach and references to the
> literature are awkward. The references to profile likelihood work
> suggest that there is a theoretical basis for the approach (that paper
> suggests consistency) but the settings considered in such work are
> quite different, I believe. Integrated likelihood usually refers to
> something different too. Finally, as Reviewer 3 makes clear the only
> current implementation of the approach was intended as an
> approximation to ML rather than as ML. The central issue is whether
> treating unobserved data (ancestral states here) as parameters
> sometimes works; the EM algorithm were developed with the idea that
> this is not what should be done. In any case, references directly tied
> to that issue would be more relevant. Additional suggestions include:
> 1) Make a clear statement in the intro about why the usual likelihood
> is considered the right likelihood: it is the probability of the
> *observed* data. 2) The point that large numbers of parameters need
> not defeat likelihood methods is valid but it should be made clear
> when referencing such work that the setting is different.

We use "integrated likelihood" in the same manner as Goldman. We have replaced "profile likelihood" with "joint objective function" throughout, unless we were specifically making a comparison with profile methods. We have removed references that are unrelated to the approaches considered in the manuscript, and defer these to the discussion for the interested reader to follow up with if they so choose. Thanks.

Thank you for your suggestion in point 1; we have added it to the introduction.
We have also clarified re point 2 that the joint objective function is related to profile likelihoods, but is not the same thing.

> -l21: Some rewording is needed. The joint method will, as n gets
> large, continue to estimate a well-resolved topology with positive
> probability; albeit with small middle edge-length.

We have removed all instances of "multifurcating" with "zero internal branch length" except for two occurrences in the Discussion.

> -l131: A reader might be able to avoid the appendix (which is
> difficult to wade into) if an example zeta in a four taxon case was
> given as illustration.

We realized that we could eliminate the xi (xi looks like zeta) from the notation in the main text. This is definitely an improvement in readability-- thanks!

> -l150: I initially found this paragraph confusing. Table S1 is a
> result that you are using as intuition, so you can't be giving
> intuition for the entire result (otherwise it gets circular). I think
> you are trying to give intuition for the `everywhere except a set of
> measure' portion of the result rather than entire result. Perhaps `the
> proof that hat t neq t* except on a set of measure 0 can be seen
> intuitively...'.

We have added the suggested wording, thanks.

> -l154 and (should check elsewhere): `the limiting value of the
> estimator' rather than `the estimator' is what is being referred to
> here.

We focus in this paper in the limiting case only.
We have added a sentence clarifying language: "Because we are interested in the properties of joint inference in the limit of $n$ going to infinity, we will focus on estimates in this case and simply refer to the right hand side of \eqref{eq:exp-log-lik} as ``the estimator.''"

> -l186: which `likelihood terms' is not that clear to me: `each site
> results in multiple factors, only one of which depends on w with
> dependence being through (1+w) or (1-w) depending ...'

We have clarified this by specifying that we're talking about the joint likelihood expression here.

> -l189: `depends on the values of other branch parameters' or some
> other rewording. I don't think it is that they are unknown. It is that
> they need optimization, so you can't optimize w independently of them.

We have adopted the suggested wording, thanks.

> -l201, Figs 2 and 3: I don't understand why direct numerical
> optimization was needed. You have the exact expressions for hat w in
> Fig S2. So you could have plotted them as Fig 2. A similar comment
> applies to Fig 3. Are you trying to convince the reader that your
> proofs have no errors? That seems more an issue for supp material.

We have rewritten this section to clarify what is being used where. Specifically, although we can find the regions for which the various ancestral state splits are optimal, we cannot obtain optimal branch lengths analytically.

We moved these figures into their current locations in response to a previous reviewer comment, who thought that Figure S2 was quite uninteresting while Figure 2 was relevant. We would prefer to leave them be. Figure 3 is an essential part of our argument that the internal branch length has substantial deviation from truth.

> -l252-259: I don't think this is an example of profile likelihood as
> considered in Murphy and Van der Vaart.

We agree, and have made the requisite changes: both replacing references to "profile likelihood" in the body of the text with "joint objective function", and by clarifying that they aren't quite the same things in the Discussion.

> -l264: Do you mean that `consistency can hold even when the numbers of
> parameters tend to infinity'? The statement `not a necessary
> condition' just means that inconsistency can arise in different ways
> (eg. lack of identifiability).

We have removed this reference in the Discussion. We use it to show a relatively well-known example where maximum likelihood fails for a similar reason as in the current case of joint inference of phylogeny. The discussion preceding this sentence shows that parameters tending to infinity are not a sufficient condition for inconsistency, which we believe is more clear having removed the reference to Neyman and Scott. We have removed the reference to the paper by LeCam. It explains various setting where maximum likelihood fails, including Neyman and Scott's example, but is not directly relevant to our argument.

> Reviewer: 2

> Comments to the Author
> Shaw et al provided a revised version of the manuscript that I
> reviewed about a year ago showing that joint estimation of ancestral
> states and branch length is in general not consistent. The authors now
> map out the regions in which topologies are incorrectly estimated and
> include a plot that shows the bias of the internal branch length for
> different tree configurations. I think this is a useful piece of work.
> But some parts are completely unsurprising, while some empirical
> results seem inaccurate or misleading.

> The most important results/claims by Shaw et al are:

> 1) "jointly estimating phylogenetic parameters along with ancestral
> states is not consistent in general"... "only correct on a set of
> measure 0".
> It is clearly true that branch length will be biased. But the fact
> that joint estimation will not recover the true branch lengths is not
> surprising but perfectly expected. Joint estimation ignores subleading
> configurations of internal states. Since the subleading terms
> typically require more state changes, branch length estimates ignoring
> them are biased. In a similar way a Taylor expansion is almost nowhere
> "correct", but still a useful and insightful approximation (a better
> analogy is ignoring looped diagram in a diagrammatic expansion.).

I'm afraid that this fact wasn't obvious to us at the outset of this project. The interest that the other reviewers have shown in the manuscript confirms that it is not broadly obvious.

> 2) In some parameter regions the internal branch of the four taxon
> tree collapses to 0.
> Again, this is true and identifying regions in parameter space where
> the internal branch collapses to zero is the most interesting part of
> the manuscript in my opinion. Unfortunately, the manuscript doesn't
> give much insight as to why the length of the internal branch shrinks
> to zero in some regions of parameter space. There is, however, a
> rather intuitive explanation for this. In the inv Fels topology, the
> internal branch collapses to 0 if
>
> p_x^2 > c p_y
>
> (the prefactor c depends on specifics like alphabet size etc). And
> this has an intuitive interpretation: if p_x^2 > p_y, it is more
> likely to observe a mutation twice on two branches of length p_x than
> once on a branch of length p_y. As a result, the joint inference will
> push p_w to zero and assign all state changes on the terminal branches
> of length p_x. But for short branches, this corresponding sector of
> parameter space is a small (tending to zero for a fixed p_x/p_y with
> p_x \to 0).

We have added this intuition, thank you!

> My empirical tests of the approximate joint inference scheme suggest
> this region is much smaller than the one identified by Shaw et al in
> Fig 3.

We have worked this out with Dr. Neher and found that the difference lay in that he was not optimizing the objective function thoroughly and was getting stuck in a local optimum.

> 3) "[The bias is] of the same order as the branch lengths themselves."
> This is incorrect or misleading, depending on how you read it. In
> typically cases, the error made by this procedure should be quadratic
> in branch length and hence of higher order than the branch length
> itself. All empirical test we have done via numerically optimizing
> branch lengths suggests this is true for typical trees and that the
> errors made are small provided ALL adjacent branch lengths are short.
> That said, relative error of one branch can be large if adjacant
> branches are long such as the case above. In this case the error is
> quadratic in the length of the longer adjacent branches rather than
> the branch itself.

We have removed the "order" language, instead just focusing on the downward bias.

> Reviewer: 3

> Comments to the Author
> This a resubmission of an earlier version of the paper, for which I
> was reviewer #4.  The other reviewers of the previous submission did
> an excellent job in pointing out minor problems and making suggestions
> for improvement.  It seems that Shaw et al. have addressed all of
> these suggestions in exemplary fashion, and  think the result is a
> much stronger paper.  I especially appreciate the authors' efforts to
> make the mathematical derivations and proofs more accessible to
> general readers (like myself).

> I often write long reviews but given the careful attention to detail
> in responding to the previous reviews of an already good paper, I have
> nothing further to add.

Thank you for your kind words.

> Reviewer: 4

> Comments to the Author
> The authors have made some substantial changes addressing the concerns
> of the last round and I think that has improved the clarity and in
> some instances also strengthened the results.
> Remaining comments are mostly about the notation which I still find a
> bit hard to follow.
> Why are site and ancestral node splits carrying this distinction of
> whether the m'th taxon (or node) has a 1?

> Perhaps it's a latexdiff leftover but eq 6 in line 151 seems to have a
> line with just a ; \xi )

Indeed, this was not present in the non-latexdiff pdf.

> Line 593 p 25 I don't understand "the mapping \psi operates by
> returning the tips labeled as 1 in a site pattern to obtain a site
> split in P(Y) if the set of tips labeled 1 is not in P(Y)" and this
> mapping seems fundamental. Perhaps you mean to refer to the m'th taxon
> (or node)..?

We have rewritten this confusing sentence in terms of complements, and have added a little bit of text pointing the reader to the formal definition just above.

> The symmetric model is assumed strongly in the first part of the
> appendix, with the factor of 2; my guess is that the results hold in
> asymmetric models but the notation is harder, is that correct?

Unfortunately a lot would have to change to generalize from the symmetric model. We would not be able to "collapse" site splits, and the Hadamard transform only works for submodels of the Kimura 3 parameter model. See, for example:

Bryant, D. (2009). Hadamard Phylogenetic Methods and the n-taxon Process. Bulletin of Mathematical Biology, 71(2), 339–351. https://doi.org/10.1007/s11538-008-9364-8

> What's q in equation 7 in the appendix (line 655) ?

We have inserted text reminding the reader that this is the number of site splits.

> More substantitvely, the limit in (8) seems not established because as
> n -> \infty so would q and n_j, so these can't be treated as fixed .

We have added text emphasizing that the number of taxa is fixed so q is fixed, and that although n_j goes to infinity the ratio n_j/n converges to the probability of seeing that column.

> page 9 - at this point in the paper, unless the reader has already
> read the appendix they don't know what the estimator means or what the
> p_ytilde j are - so this text doesn't make sense at this point.

This is a good point! We have rephrased things, eliminating p_ytilde_j except for a parenthetical reference to the Appendix. We believe that the result is much more readable.

> page 30 line 667-8 what are S and Y and what makes a path? I gather
> from the math that you mean paths in the tree, since the sum is over
> ancestral states and there's a product of probabilities of changes on
> branches but (perhaps my lack of background) the form of (9) isn't
> obvious I think because I am not exactly clear on what the notation
> means.

We have clarified the definitions of S and Y, and added that by path we mean graph-theoretic path. We are afraid that adding more detail about the Hadamard transform is beyond the scope of this paper, and it is well described in Semple and Steel's book.